---
title: "Capsid position correlation"
output:
  html_document:
    keep_md: yes
---

Report generated at: `r format(as.POSIXct(Sys.time(), format = "%y%m%d %H:%M"), "%m/%d/%Y %I:%M %p")` PDT

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(gridExtra)
library(plotly)
library(ggbeeswarm)
library(viridis)
library(ggiraphExtra) 
library(Rfit)


knitr::opts_chunk$set(
  echo = FALSE,
  dev = c("png", "svg")
)

apply_theme <- function(p) {
  p = p + theme(
      # legend.position = "none",
      # legend.title = element_blank(),
      # axis.title.x = element_blank(),
      # axis.title.y = element_blank(),
      # axis.text.x = element_blank(),
      axis.title.y = element_text(size=10),
      axis.title.x = element_text(size=10),
      # plot.title = element_text(hjust = 0.5, size=12),
      strip.background = element_blank(),
      strip.text = element_blank(),
      panel.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
    )
  p
}


```

## Meaning of correlation_data.csv

- pos: all rows are calculated by each postion of capsid
- Entropy: entropy of mutations >= 0.1%
- Subtype Chi-square test was calculated by contigency table of subtypes and mutations >= 0.1% in any subtype.
  - Subtype Chi-square
  - Subtype Chi-square rank: Subtype chi-square rank number
- HLA References: Number of HLA References
- HLA Chi-square was calculated by the contigency table of mutations and HLA coexist or not.
  - HLA Chi-square sum: sum of HLA Chi-square test by position
  - HLA Chi-square median: median of HLA Chi-square test by position
  - HLA Chi-square mean: mean of HLA Chi-square test by position


## Subtype chi-square test rank and #HLA Reference, regression
```{r sct-rank-hla}
df = read.csv('correlation_data.csv') %>% mutate(subtype_rank=Subtype.Chi.square_rank, HLA=HLA.References)

model = lm(Entropy ~ subtype_rank + HLA, data=df)
summary(model)
confint(model)

# ggPredict(model,se=TRUE,interactive=TRUE)
```

## Subtype chi-square test rank and #HLA Reference with interaction, regression
```{r sct-rank-hla-inter}
df = read.csv('correlation_data.csv') %>% mutate(subtype_rank=Subtype.Chi.square_rank, HLA=HLA.References)

model = lm(Entropy ~ subtype_rank * HLA, data=df)
summary(model)
confint(model)

# ggPredict(model,se=TRUE,interactive=TRUE)
```

## Subtype chi-square test and #HLA Reference, Rfit
```{r sct-hla}
df = read.csv('correlation_data.csv') %>% mutate(subtype=Subtype.Chi.square, HLA=HLA.References)

model = rfit(Entropy ~ subtype + HLA, data=df)
summary(model)
confint(model)

df = df %>% filter(Entropy > 0)
spearman = cor.test(df$Entropy, df$subtype, method = "spearman")
print("Entropy x Subtype chi-square test")
print(spearman)

df = df %>% filter(Entropy > 0)
spearman = cor.test(df$Entropy, df$HLA, method = "spearman")
print("Entropy x #HLA Reference")
print(spearman)

df = df %>% filter(Entropy > 0)
spearman = cor.test(df$subtype, df$HLA, method = "spearman")
print("Subtype chi-square test x #HLA Reference")
print(spearman)

```

## Subtype chi-square test and #HLA Reference with interaction, Rfit
```{r sct-hla-inter}
df = read.csv('correlation_data.csv') %>% mutate(subtype=Subtype.Chi.square, HLA=HLA.References)

model = rfit(Entropy ~ subtype * HLA, data=df)
summary(model)
confint(model)
```

## Subtype chi-square test and HLA chi-square test mean, Rfit

Positions without HLA Chi-square test were treated as 0.

```{r sct-hct-mean}
df = read.csv('correlation_data.csv') %>% 
  mutate(subtype=Subtype.Chi.square, HLA=HLA.Chi.square.mean) %>%
  mutate(HLA=replace_na(HLA, 0))

model = rfit(Entropy ~ subtype + HLA, data=df)
summary(model)
confint(model)

df = df %>% filter(Entropy > 0) 
spearman = cor.test(df$Entropy, df$HLA, method = "spearman")
print("Entropy x HLA chi-square test mean")
print(spearman)
```

## Subtype chi-square test and HLA chi-square test mean with interaction, Rfit

Positions without HLA Chi-square test were treated as 0.

```{r sct-hct-mean-inter}
df = read.csv('correlation_data.csv') %>% 
  mutate(subtype=Subtype.Chi.square, HLA=HLA.Chi.square.mean) %>%
  mutate(HLA=replace_na(HLA, 0))

model = rfit(Entropy ~ subtype * HLA, data=df)
summary(model)
confint(model)
```

## Subtype chi-square test and HLA chi-square test sum, Rfit

Positions without HLA Chi-square test were treated as 0.

```{r sct-hct-sum}
df = read.csv('correlation_data.csv') %>% 
  mutate(subtype=Subtype.Chi.square, HLA=HLA.Chi.square.sum) %>%
  mutate(HLA=replace_na(HLA, 0))

model = rfit(Entropy ~ subtype + HLA, data=df)
summary(model)
confint(model)

df = df %>% filter(Entropy > 0)
spearman = cor.test(df$Entropy, df$HLA, method = "spearman")
print("Entropy x HLA chi-square test sum")
print(spearman)
```

## Subtype chi-square test and HLA chi-square test sum with interaction, Rfit

Positions without HLA Chi-square test were treated as 0.

```{r sct-hct-sum-inter}
df = read.csv('correlation_data.csv') %>% 
  mutate(subtype=Subtype.Chi.square, HLA=HLA.Chi.square.sum) %>%
  mutate(HLA=replace_na(HLA, 0))

model = rfit(Entropy ~ subtype * HLA, data=df)
summary(model)
confint(model)
```


## Subtype chi-square test and HLA chi-square test median, Rfit

Positions without HLA Chi-square test were treated as 0.

```{r sct-hct-median}
df = read.csv('correlation_data.csv') %>% 
  mutate(subtype=Subtype.Chi.square, HLA=HLA.Chi.square.median) %>%
  mutate(HLA=replace_na(HLA, 0))

model = rfit(Entropy ~ subtype + HLA, data=df)
summary(model)
confint(model)

df = df %>% filter(Entropy > 0)
spearman = cor.test(df$Entropy, df$HLA, method = "spearman")
print("Entropy x HLA chi-square test median")
print(spearman)
```

## Subtype chi-square test and HLA chi-square test median with interaction, Rfit

Positions without HLA Chi-square test were treated as 0.

```{r sct-hct-median-inter}
df = read.csv('correlation_data.csv') %>% 
  mutate(subtype=Subtype.Chi.square, HLA=HLA.Chi.square.median) %>%
  mutate(HLA=replace_na(HLA, 0))

model = rfit(Entropy ~ subtype * HLA, data=df)
summary(model)
confint(model)
```